### 2015-02-20 sicp/Cohort 1/chapter-1.2/ndpar-1.2.tex
Just doing a bit of housekeeping here
I don't think this lost anything but some of the deletions are a
little concerning because they should have a corresponding creation.

\documentclass[12pt]{article}

\usepackage{amsthm}
\usepackage{amsmath}    % need for subequations
\usepackage{amsfonts}
\usepackage[ocgcolorlinks]{hyperref}

\setlength{\baselineskip}{16.0pt}    % 16 pt usual spacing between lines

\setlength{\parskip}{3pt plus 2pt}
\setlength{\parindent}{20pt}
\setlength{\oddsidemargin}{0.5cm}
\setlength{\evensidemargin}{0.5cm}
\setlength{\marginparsep}{0.75cm}
\setlength{\marginparwidth}{2.0cm}
\setlength{\marginparpush}{1.0cm}
\setlength{\textwidth}{150mm}

\DeclareMathOperator{\sinc}{sinc}
\DeclareMathOperator{\sgn}{sgn}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}

\newenvironment{definition}[1][Definition]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
\newenvironment{example}[1][Example]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
\newenvironment{remark}[1][Remark]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}

\begin{document}

\section{Example: Counting change}

\begin{lemma}
    $\forall n, d, k \in \mathbb{N} : d, k \ge 1 \implies n^{k-1} + (n-d)^k \le n^k$
\end{lemma}

\begin{proof}
    $$d \ge 1 \implies n - d \le n - 1 < n.$$
    $$k \ge 1 \implies (n-d)^{k-1} < n^{k-1}.$$
    $$(n-d)^k = (n-d)(n-d)^{k-1} \le (n-1)n^{k-1} = n^k - n^{k-1}.$$
\end{proof}

\begin{definition}
    $T(n,k)$ is the number of ways to change amount $n$ using $k$ kinds of coins. $n, k \in \mathbb{N}, n, k \ge 1$.
\end{definition}

\begin{theorem}
    $T(n,k) = O(n^k).$
\end{theorem}

\begin{proof}
    To prove the theorem we have to show that $\exists c>0$ such that $T(n,k) \le cn^k$ for sufficiently large $n$.
    Let's prove it using substitution method (CLRS, p.83).

    Assume that the inequality is true.
    The recurrence formula for $T(n,k)$ is
    $$T(n,k) = T(n,k-1) + T(n-d,k),$$
    where $d$ is denomination of the first coin. $d \ge 1$. Therefore by Lemma
    $$T(n,k) \le cn^{k-1} + c(n-d)^k = c\left[n^{k-1} + (n-d)^k\right] \le cn^k.$$

    To finish the proof, we just need to find $c$. Since the boundary condition $T(1,1) = 1$, it's sufficient to put $c \ge 1$.
\end{proof}

\end{document}