### 2016-01-29 pfpl/cheatsheet.tex
Add Chapter 2

\documentclass[12pt]{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{proof}
\usepackage{hyperref}

\usepackage[margin=.25in]{geometry}
\begin{document}
\begin{tabular}[b] {p{.05\textwidth} p{.13\textwidth} p{.15\textwidth} p{.22\textwidth} p{.32\textwidth}}
Page & Expression & Name & Say & Meaning \\ \\
5& $\mathcal{P}$ & Script P & Proposition & Something to be proved \\
5& $\mathcal{P}(\it{a})$ & Script P of $\it{a}$ & Proposition about tree $\it{a}$ & Something to be proved about AST $\it{a}$ \\
5& $\mathcal{O}$ & Script O & Operator & An operator that can be used in an AST \\
5& $\mathcal{O}(\it{a})$ & Script O of $\it{a}$ & Operator of arity $\it{a}$ & An operator of a given arity \\
5& $\mathcal{X}_{\it{s}}$ & Script X sub s & Variables $\it{x}$ of sort $\it{s}$ & Variables $\it{x}$ of sort $\it{s}$ \\
5& $\it{S} $ & S & A set of sorts & A set of sorts \\
5& $ \{ \it{X}_{s} \} _{\it{s} \in \mathcal{S}}$  & Family & Family $ \it{X} $ of $\it{s}$ & a sort-indexed family of disjoint finite sets $ \it{X}_s $ of variables $\it{x}$ of sort $ \it{s} $\\
6& $\left[ \it{b} / \it{x}\right] \it{a}$ & Substitution & Substitute $\it{b}$ for $\it{x}$ in $\it{a}$ & Substitute  $\it{b}$ for $\it{x}$ in $\it{a}$ \\
7& $ \it{x}_1, ... ,\it{x}_{\it{n}}.\it{a} $ & Abstractor & Bind variables $\it{x_n}$ to expression $\it{a}$ & Bind variables $\it{x_n}$ to expression $\it{a}$ \\
8& $ \overrightarrow{\it{x}} $ & X arrow & List of $\it{x}s$ & $ \it{x}_1, ... ,\it{x}_{\it{n}} $ \\
8& $ \rho : \overrightarrow{\it{x}} \leftrightarrow \overrightarrow{\it{x}}'$ & Fresh renaming & Freshen $ \it{x} $ using renaming $ \rho $ & A bijection between $ \overrightarrow{\it{x}} $ and $ \overrightarrow{\it{x}}' $ where $ \overrightarrow{\it{x}}'$ is fresh. \\
8& $\widehat{\rho}_{\it{i}} (a_{\it{i}})$ & Rho hat sub i & Rename result & The result of applying the renaming $ \rho_ {\it{i}} $ to $ \it{a}_{\it{i}} $ \\
8& $x =_\alpha y$ & Equal alpha & $ \it{\alpha}$-$\it{equivalence} $ & Trees x and y equal up to renaming \\
9& $x \overset{\Delta}{=} y$ & Delta equals & Definitially equals & Trees x and y are definitionally equal \\
13& $\tau$ type& type & type $\tau$ & Judgement that $\tau$ is a type \\
13& $\it{e} : \tau$ & expression type & expression $\it{e}$ is of type $\tau$ & Judgement that expression $\it{e}$ is of type $\tau$ \\
13& $\it{e} \Downarrow \it{v}$ & expression value & expression $\it{e}$ has value $\it{v} $ & Judgement that expression $\it{e}$ has value $\it{v} $\\
14& $\infer {\it{J}} {\it{J}_{1}...\it{J}_{\it{k}}}$  & surfboard & infers & Judgements ${\it{J}_{1}...\it{J}_{\it{k}}}$ infer judgement ${\it{J}}$\\

\end{tabular}

\end{document}