### 2016-05-13 pfpl/cheatsheet.tex
Update cheat sheet

\documentclass[12pt]{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{proof}
\usepackage{hyperref}

\usepackage[margin=.25in]{geometry}
\begin{document}
\title{Cheat Sheet for PFPL}
\maketitle
\subsection*{I. Judgements and Rules}
\subsubsection*{1 Abstract Syntax}
\begin{tabular}[b] {p{.025\textwidth} p{.025\textwidth} p{.13\textwidth} p{.15\textwidth} 
p{.22\textwidth} p{.32\textwidth}}
N & P & Expression & Name & Say & Meaning \\ \\
1& 5& $\mathcal{P}$ & Script P & Proposition & Something to be proved \\
2& 5& $\mathcal{P}(\it{a})$ & Script P of $\it{a}$ & Proposition about tree $\it{a}$ & Something to be proved about AST $\it{a}$ \\
3& 5& $\mathcal{O}$ & Script O & Operator & An operator that can be used in an AST \\
4& 5& $\mathcal{O}(\it{a})$ & Script O of $\it{a}$ & Operator of arity $\it{a}$ & An operator of a given arity \\
5& 5& $\mathcal{X}_{\it{s}}$ & Script X sub s & Variables $\it{x}$ of sort $\it{s}$ & Variables $\it{x}$ of sort $\it{s}$ \\
6& 5& $\it{S} $ & S & A set of sorts & A set of sorts \\
7& 5& $ \{ \it{X}_{s} \} _{\it{s} \in \mathcal{S}}$  & Family & Family $ \it{X} $ of $\it{s}$ & A sort-indexed family of disjoint finite sets $ \it{X}_s $ of variables $\it{x}$ of sort $ \it{s} $\\
8& 6& $\left[ \it{b} / \it{x}\right] \it{a}$ & Substitution & Substitute $\it{b}$ for $\it{x}$ in $\it{a}$ & Substitute  $\it{b}$ for $\it{x}$ in $\it{a}$ \\
9& 7& $ \it{x}_1, ... ,\it{x}_{\it{n}}.\it{a} $ & Abstractor & Bind variables $\it{x_n}$ to expression $\it{a}$ & Bind variables $\it{x_n}$ to expression $\it{a}$ \\
10& 8& $ \overrightarrow{\it{x}} $ & X arrow & List of $\it{x}s$ & $ \it{x}_1, ... ,\it{x}_{\it{n}} $ \\
11& 8& $ \rho : \overrightarrow{\it{x}} \leftrightarrow \overrightarrow{\it{x}}'$ & Fresh renaming & Freshen $ \it{x} $ using renaming $ \rho $ & A bijection between $ \overrightarrow{\it{x}} $ and $ \overrightarrow{\it{x}}' $ where $ \overrightarrow{\it{x}}'$ is fresh. \\
12& 8& $\widehat{\rho}_{\it{i}} (a_{\it{i}})$ & Rho hat sub i & Rename result & The result of applying the renaming $ \rho_ {\it{i}} $ to $ \it{a}_{\it{i}} $ \\
13& 8& $x =_\alpha y$ & Equal alpha & $ \it{\alpha}$-$\it{equivalence} $ & Trees x and y equal up to renaming \\
14& 9& $x \overset{\Delta}{=} y$ & Delta equals & Replacement & Replace expression x with expression y \\
\end{tabular}

\subsubsection*{2 Inductive Definitions}
\begin{tabular}[b] {p{.025\textwidth} p{.025\textwidth} p{.13\textwidth} p{.15\textwidth} 
p{.22\textwidth} p{.32\textwidth}}
N & P & Expression & Name & Say & Meaning \\ \\

15& 13& $\tau$ type& Type & Type $\tau$ & Judgement that $\tau$ is a type \\
16& 13& $\it{e} : \tau$ & Colon &  $\it{e}$ is of type $\tau$ & Judgement that expression $\it{e}$ is of type $\tau$ \\
17& 13& $\it{e} \Downarrow \it{v}$ & Down arrow &  $\it{e}$ has value $\it{v} $ & Judgement that expression $\it{e}$ has value $\it{v} $\\
18& 14& $\infer {\it{J}} {\it{J}_{1}...\it{J}_{\it{k}}}$  & Surfboard & Infers & Judgements ${\it{J}_{1}...\it{J}_{\it{k}}}$ infer judgement ${\it{J}}$\\

\end{tabular}

\subsubsection*{3 Hypothetical and General Judgements}
\begin{tabular}[b] {p{.025\textwidth} p{.025\textwidth} p{.13\textwidth} p{.15\textwidth} 
p{.22\textwidth} p{.32\textwidth}}
N & P & Expression & Name & Say & Meaning \\ \\
19& 23 & ${\it{J}_{1}...\it{J}_{\it{k}}} \vdash_{\mathcal{R}} \mathcal{K}$ & Turnstile & Entails & Given $\mathcal{R}$ and $\it{J}$ infer $\mathcal{K}$\\

20& 23 & $\Gamma$ & Gamma & Judgements Gamma & A finite set of judgements \\
21& 23 & $\Delta$ & Delta & Judgements Delta & A finite set of type judgements \\
22& 25 & $\Gamma \models_{\it{R}} \it{J}$ & Double turnstile & Admissible & $\vdash_{\it{R}} \Gamma$ implies $\vdash_{\it{R}} \it{J}$ \\
23& 28 & $\bigtriangledown$ & Down triangle & Generic derivation & Generic derivation \\
\end{tabular}

\subsection*{II. Statics and Dynamics}

\subsubsection*{4 Statics}
\begin{tabular}[b] {p{.025\textwidth} p{.025\textwidth} p{.13\textwidth} p{.15\textwidth} 
p{.22\textwidth} p{.32\textwidth}}
N & P & Expression & Name & Say & Meaning \\ \\

24& 36 & $ \it{n} ::= \it{s} $ & Colon colon equals & The syntax of $\it{n}$ is $\it{s}$ & Specifies the syntax of $\it{n}$ \\
25& 36 & $;$ & Semicolon & And & Separates arguments to expressionsin abstact notation\\

\end{tabular}

\subsubsection*{5 Dynamics}
\begin{tabular}[b] {p{.025\textwidth} p{.025\textwidth} p{.13\textwidth} p{.15\textwidth} 
p{.22\textwidth} p{.32\textwidth}}
N & P & Expression & Name & Say & Meaning \\ \\
26& 41 & $\it{s} \longmapsto \it{s'}$ & Bar arrow & Transistion  & State $\it{s}$ transitions to state $\it{s'}$ \\

27& 42 & $\it{s} \longmapsto ^{*} \it{s'}$ & Bar arrow star & Iterated transistion  & State $\it{s}$ transitions to state $\it{s'}$ over more than zero transitions \\

28& 42 & $\it{s} \longmapsto ^{n} \it{s'}$ & Bar arrow n & N times iterated transistion  & State $\it{s}$ transitions to state $\it{s'}$ over n transitions \\

29& 44 & $\mathcal{E}$  & Script E & Expression context & Expression context \\

30& 45 & $\circ $ & Circle & Hole & Placeholder to put an instruction \\
31& 46 & $ \it{e} \equiv \it{e}' $ & Equivalent & Definitional equivalence & $\it{e}$ is definitionally equivalent to $\it{e}'$ \\

\end{tabular}

\subsubsection*{6 Type Safety}
\begin{tabular}[b] {p{.025\textwidth} p{.025\textwidth} p{.13\textwidth} p{.15\textwidth} 
p{.22\textwidth} p{.32\textwidth}}
N & P & Expression & Name & Say & Meaning \\ \\
32& 58 & $\it{e}??$ & Wrong & E goes wrong & Expression $\it{e}$ goes wrong \\

\end{tabular}
\subsubsection*{7 Evaluation Dynamics}

\begin{tabular}[b] {p{.025\textwidth} p{.025\textwidth} p{.13\textwidth} p{.15\textwidth} 
p{.22\textwidth} p{.32\textwidth}}
N & P & Expression & Name & Say & Meaning \\ \\

33& 58 & $\it{e} \Downarrow ^{\it{k}}\it{v}$  & Downarrow k & E evaluates in k steps & Expression $ \it{e}$ evaluates to $\it{v}$ in $\it{k}$ steps\\
\end{tabular}

\subsection*{III. Total Functions}
\subsubsection*{8 Function Definitions and Values}

\begin{tabular}[b] {p{.025\textwidth} p{.025\textwidth} p{.13\textwidth} p{.15\textwidth} 
p{.22\textwidth} p{.32\textwidth}}
N & P & Expression & Name & Say & Meaning \\ \\

34& 63 & $\{\it{f}\}$ & Brace brackets & Function & Surround function $\it{f}$ in abstract notation \\

35& 63 & $\it{f}.\it{e}$ & Dot & Dot & Introduces the scope $\it{e}$ of a function $\it{f}$ in abstract notation \\
36& 64 & $\it{f}(\tau_1):\tau_2$ & Function & Function definition & A function taking an argument of type $\tau_1$ and returning a value of type $\tau_2$ \\

37& 64 &  $[\![x.e/f]\!]e'$ & Script bracket & Function substitution & Function substitution \\

38& 65 & $\tau_1 \to \tau_2$ & Right arrow & Maps to & A total function that maps elements of type $\tau_1$ to elements of type $\tau_2$ \\
39& 65 & $\lambda $ & Lambda & Lambda & Abstraction \\
\end{tabular}
\subsubsection*{9 System T of Higher-Order Recursion}

\begin{tabular}[b] {p{.025\textwidth} p{.025\textwidth} p{.13\textwidth} p{.15\textwidth} 
p{.22\textwidth} p{.32\textwidth}}
N & P & Expression & Name & Say & Meaning \\ \\

40& 71 & $\hookrightarrow$ & Hook arrow & Select & Selector (used in System T recursion, sum types, and product types)\\
41& 71 & $|$ & Bar & Either & A choice \\
42& 71 & $\overline{\it{n}}$ & Overline & Church numbering & The Church numbering \\
43& 76 & $\lceil n \rceil$ & Divided hat & G$\ddot{o}$del numbering & The G$\ddot{o}$del numbering \\
\end{tabular}
\subsection*{IV. Finite Data Types}
\subsubsection*{10 Product Types}
\begin{tabular}[b] {p{.025\textwidth} p{.025\textwidth} p{.13\textwidth} p{.15\textwidth} 
p{.22\textwidth} p{.32\textwidth}}
N & P & Expression & Name & Say & Meaning \\ \\

44& 81 & $<>$ & Angle brackets & Null tuple & Null tuple \\
45& 81 & $<e_1,e_2>$ & Angle brackets & Ordered pair & Ordered pair \\
46& 81 & $e.$l & Left & Left projection & Select left member of the ordered pair \\
47& 81 & $e.$r & Right & Right projection & Select right member of the ordered pair \\
\end{tabular}
\subsubsection*{11 Sum Types}
\begin{tabular}[b] {p{.025\textwidth} p{.025\textwidth} p{.13\textwidth} p{.15\textwidth} 
p{.22\textwidth} p{.32\textwidth}}
N & P & Expression & Name & Say & Meaning \\ \\

48& 87 & l.$e$ & Left & Left injection & Create sum type element using left type \\
49& 87 & r.$e$ & Right & Right injection & Create sum type element using right type \\
50& 93 & $\triangleq$ & Delta equals & Delta equals & Replacement \\
\end{tabular}

\subsection*{VI. Infinite Data Types}
\subsubsection*{14 Generic Programming}
\begin{tabular}[b] {p{.025\textwidth} p{.025\textwidth} p{.13\textwidth} p{.15\textwidth} 
p{.22\textwidth} p{.32\textwidth}}
N & P & Expression & Name & Say & Meaning \\ \\

51& 121 & $t.\tau$ & Dot & Type operator & Bind $t$ to type $\tau$ \\
\end{tabular}
\subsubsection*{15 Inductive and Coinductive Types}
\begin{tabular}[b] {p{.025\textwidth} p{.025\textwidth} p{.13\textwidth} p{.15\textwidth} 
p{.22\textwidth} p{.32\textwidth}}
N & P & Expression & Name & Say & Meaning \\ \\

52& 133 & $ \cong $ & Tilde equal & Isomorphism & Isomorphism \\
\end{tabular}
\subsection*{VII. Variable Types}
\subsubsection*{17 Abstract Types}
\begin{tabular}[b] {p{.025\textwidth} p{.025\textwidth} p{.13\textwidth} p{.15\textwidth} 
p{.22\textwidth} p{.32\textwidth}}
N & P & Expression & Name & Say & Meaning \\ \\
53 & 149 & $\exists(t.\tau)$ & Existential quantifier & Exists & Defines an interface \\
54 & 153 & $\forall(t.\tau\longrightarrow\tau_2)$ & Universal quantifier & For all & Defines universal type \\
\end{tabular}

\subsubsection*{18 Higher Kinds}
\begin{tabular}[b] {p{.025\textwidth} p{.025\textwidth} p{.13\textwidth} p{.15\textwidth} 
p{.22\textwidth} p{.32\textwidth}}
N & P & Expression & Name & Say & Meaning \\ \\

55& 157 & $::$ & Colon colon & Kind type constructor & Maps types to types \\
\end{tabular}
\subsection*{VIII. Partiality and Recursive Types}
\subsubsection*{19 System PCF of Recursive Functions}
\begin{tabular}[b] {p{.025\textwidth} p{.025\textwidth} p{.13\textwidth} p{.15\textwidth} 
p{.22\textwidth} p{.32\textwidth}}
N & P & Expression & Name & Say & Meaning \\ \\

56& 166 & $\mapsto$ & Short bar arrow & Maps to & Function definition \\
57& 166 &$ \bot$ & Bottom & Bottom & Totally undefined partial function \\
58& 167 & $\tau_1 \rightharpoonup \tau_2$ & Harpoon & Partial function & Partial function \\
\end{tabular}

\subsubsection*{20 System FPC of Recursive Types}
\begin{tabular}[b] {p{.025\textwidth} p{.025\textwidth} p{.13\textwidth} p{.15\textwidth} 
p{.22\textwidth} p{.32\textwidth}}
N & P & Expression & Name & Say & Meaning \\ \\

59& 177 & $\_$ & Underscore & Underscore & Unfree variable \\
\end{tabular}
\subsection*{IX. Dynamic Types}
\subsubsection*{21 The Untyped $\lambda$-Calculus}
\begin{tabular}[b] {p{.025\textwidth} p{.025\textwidth} p{.13\textwidth} p{.15\textwidth} 
p{.22\textwidth} p{.32\textwidth}}
N & P & Expression & Name & Say & Meaning \\ \\

60& 185 & $\Lambda$ & Lambda & Lambda calculus & The lambda calculus \\
61& 188 & Y & Y & Y Combinator & The Y combinator \\
62& 190 & $x^{\dagger}$ & Superscript cross & Superscript cross & Language isomorphism \\

\end{tabular}

\subsection*{X. Subtyping}
\subsubsection*{24 Structural Subtyping}
\begin{tabular}[b] {p{.025\textwidth} p{.025\textwidth} p{.13\textwidth} p{.15\textwidth} 
p{.22\textwidth} p{.32\textwidth}}
N & P & Expression & Name & Say & Meaning \\ \\

63& 213 & $\tau'<:\tau$ & Subtype & $\tau'$ is a subtype of $\tau$ & $\tau'$ is a subtype of $\tau$ \\

\end{tabular}

\subsection*{XI. Dynamic Dispatch}
\subsubsection*{27 Inheritance}
\begin{tabular}[b] {p{.025\textwidth} p{.025\textwidth} p{.13\textwidth} p{.15\textwidth} 
p{.22\textwidth} p{.32\textwidth}}
N & P & Expression & Name & Say & Meaning \\ \\

64& 252 & $()^{\ddagger}$ & Isomorphism & Isomorphism & $()^{\ddagger}$ is a method isomorphism \\

\end{tabular}
\end{document}